Illuminating a Dark Lens : A Type Ia Supernova \\ Magnified by Galaxy Cluster Abell 2744

### 2015-04-15 TABLE/magnifications.tex
updated some figures, reset the changetext coloring, updated the spectroscopy a bit

Bradac(v2)   & 2.21     & 2.23  &   2.20$-$2.28 &   \citealt{Wang:2015} & \change{Updated version of the {\tt SWUnited} model with new strong-lensing constraints from HFF imaging. Errors from bootstrap resampling only strong-lensing constraints.}\\

@@

\tablenotetext{j}{{\tt WSLAP+} : Weak and Strong Lensing Analysis Package plus member galaxies \change{(Note: no weak-lensing constraints used for Abell 2744)}}

>>>>

Bradac(v2)   & 2.21     & 2.23  &   2.20$-$2.28 &   \citealt{Wang:2015} & Updated version of the {\tt SWUnited} model with new strong-lensing constraints from HFF imaging. Errors from bootstrap resampling only strong-lensing constraints.\\

@@

\tablenotetext{j}{{\tt WSLAP+} : Weak and Strong Lensing Analysis Package plus member galaxies (Note: no weak-lensing constraints used for Abell 2744)}

### 2015-04-15 snTomas.tex
cleaned up author ref macros and preamble, updated code   * adjusting figures to reflect the new redshift from the host galaxy
updated some figures, reset the changetext coloring, updated the spectroscopy a bit

\begin{abstract}
SN \tomas\ is a Type Ia Supernova (\SNIa) discovered at
$z=1.31\pm0.01$ behind the galaxy cluster Abell 2744 ($z=0.308$). This
SN has a projected separation from the cluster core of
$\sim$40\arcsec, closer to the critical lensing line than any previous
cluster-lensed \SNIa.  As such, it provides a valuable opportunity to
confront gravitational lens models of galaxy clusters with the direct
measurement of an absolute magnification at the edge of the
strong-lensing region.  We derive a tightly constrained measure of the
peak apparent magnitude, corrected for light curve shape and
exinction.  In a cosmology-independent analysis, we find that \tomas\
is \change{$0.75\pm0.15$} magnitudes brighter than unlensed \SNeIa\ at
similar redshift, implying a lensing magnification
of \change{$\mu_{\rm obs}=2.00\pm0.28$}.  We compare this direct
measurement of the magnification against predicted magnifications
from \change{14 lens models,} representing a broad range of
methodologies and incorporating a variety of strong- and weak-lensing
constraints.  The models are collectively fairly accurate: \change{
half of the models deliver median magnifications that are consistent
with the measured $\mu$ to within the 1$\sigma$ uncertainties.}
However, the models are systematically biased to higher values: none
of the tested models predict a magnification equal to or lower than
the observed $\mu$, and three models are $2-4\sigma$ above the
measured magnification.  We evaluate possible causes for this bias,
\changeG{considering both modeling errors and misinterpretation of the SN
data. Although we find no single explanation that accounts for the
entire discrepancy, we do find that the subset of ``free-form'' lens
models are more accurate and provide more realistic uncertainties than
``parametric models'' for this sight-line.} A larger sample of 10-100
lensed \SNeIa\ could provide a robust evaluation of systematic biases
in lens models.  Until such a sample exists, it is advisable
to \change{increase the estimated uncertainty of predicted
magnifications to match the dispersion observed across multiple lens
models.}
\end{abstract}

\keywords{ supernovae: general }

@@

\changeG{
Galaxy clusters can be used as cosmic telescopes to magnify distant
background objects through gravitational lensing, which can
substantially increase the reach of deep imaging surveys.  The lensing
magnification enables the study of objects that would otherwise be
unobservable because they are either intrinsically
faint \citep[e.g.][]{Schenker:2012,Alavi:2014} or extremely
distant \citep[e.g.][]{Franx:1997,Ellis:2001,Hu:2002,Kneib:2004,Richard:2006,Richard:2008,Bouwens:2009a,Maizy:2010,Zheng:2012,Coe:2013,Bouwens:2014,Zitrin:2014b}.
Background galaxies are also {\it spatially} magnified, allowing for
studies of the internal structure of galaxies in the early universe
with resolutions of $\sim$100
pc \citep[e.g.]{Stark:2008,Jones:2010,Yuan:2011,Livermore:2015}.
}

\changeG{
Gravitational lensing can also provide a powerful window onto the {\it
transient} sky through an appropriately cadenced imaging survey.  The
flux magnification from strong-lensing clusters is especially valuable
for the study of $z>1.5$ supernovae
(SNe) \citep[e.g.]{Kovner:1988,Kolatt:1998,Sullivan:2000,Saini:2000,Gunnarsson:2003,Goobar:2009,Postman:2012},
which are still extremely difficult to characterize in unlensed
fields \citep[e.g.]{Riess:2001,Riess:2007,Suzuki:2012,Rodney:2012,Rubin:2013,Jones:2013}.
}

\changeG{
In the case of lensed SNe, we can also reverse the experimental setup:
instead of using strong-lensing clusters to study distant SNe, we can
use the SNe as tools for examining the lenses \citep{Riehm:2011}.  The
most valuable transients for testing and improving cluster lens models
would be {\it strongly lensed} SNe that are resolved into multiple
images \citep{Holz:2001,Oguri:2003}.  Transients that are lensed into
multiple images can also become cosmological tools, as the measurement
of time delays between the images can provide cosmographic information
to constrain the Hubble parameter \citep{Refsdal:1964} and other
cosmological parameters \citep{Linder:2011}.  We have recently
observed the first example of a multiply-imaged
SN \citep{Kelly:2015}, and expect to detect the reappearance of this
object (called ``SN Refsdal'') within 5 years, delivering a precise
test of lens model predictions \citep{Oguri:2015,Sharon:2015}.
Although detections of such objects are currently very
unlikely \citep{Li:2012}, they will become much more common in the
next decade \citep{Coe:2009,Dobke:2009}, and may be developed into an
important new cosmological tool.
}

\changeG{
In addition to time delays from multiply-imaged SNe, we can also
put cluster mass models to the test with the much more common category
of lensed Type Ia SNe.  
\citet[][hereafter P14]{Patel:2014} and \citet{Nordin:2014} presented
independent analyses of three lensed SNe, of which at least 2 are
securely classified as Type Ia SNe -- all found in the Cluster Lensing
and Supernova survey with Hubble (CLASH, PI:Postman, HST Program ID
12068, \citealt{Postman:2012}).  Both groups demonstrated that these
standard candles can be used to provide accurate and precise
measurements of the true absolute magnification along a random sight
line through the cluster.  Although in these cases the SNe were used
to {\it test} the cluster mass models, one could in principle
incorporate the measured magnifications of Type Ia SNe into the
cluster as additional model constraints.  In that role, Type Ia SNe
have the particular value that they can be found anywhere in the
cluster field.  Thus, they can deliver model constraints in regions of
``middle distance'' from the cluster core, where both strong- and
weak-lensing constraints are unavailable.
}

\change{
One of the key values in observing standard candles behind
gravitational lenses is \changeG{in addressing the problem of the
mass-sheet degeneracy} \citep{Falco:1985,Bradac:2004}.  This
degeneracy arises because one can introduce into a lens model an
unassociated sheet of uniform mass in front of or behind the
lens, without disturbing the primary observable quantities.  For
example, take a lens model with a given surface mass density $\kappa$,
and then transform the surface mass density to
$\kappa^{\prime}=(1-\lambda)\kappa+\lambda$ for any arbitrary value
$\lambda$.  Both the $\kappa$ and $\kappa^{\prime}$ models will
produce exactly the same values for all positional and shear
constraints from strong and weak lensing.  When lensed background
sources are available across a wide range of redshifts (as is the case
for the Abell 2744 cluster discussed here), it should in principle be
possible to account for the simplest form of a single mass sheet
within that redshift range.  However, there are more complex versions
of positional constraint
degeneracies \citep{Schneider:2014,Liesenborgs:2012}.  Such
degeneracies do not extend to the {\it absolute} magnification of a
background source's flux and size.  Therefore, an absolute measurement
of magnification from a standard candle or a standard ruler
\citep{Sonnenfeld:2011}  can break these fundamental degeneracies
\citep{Holz:2001}.
}

In Section~\ref{sec:DiscoveryAndFollowup} we present the discovery and
follow-up observations of SN \tomas\ at $z=1.31$, discovered behind the
galaxy cluster Abell 2744.  Sections~\ref{sec:Spectroscopy}
and \ref{sec:PhotometricClassification} describe the spectroscopy and
photometry of this SN, leading to a classification of the object as a
normal Type Ia SN.  In Section~\ref{sec:LightCurveFitting} we make a
direct measurement of the magnification of this source due to
gravitational lensing, and compare to predictions from lens models.
Finally, Section~\ref{sec:Discussion} discusses the tension between
our magnification measurement and the lens models, with implications
for the systematic error budget in magnification estimates for other
high-$z$ lensed sources.

@@

SN \tomas\ was discovered in Hubble Space Telescop (\HST) observations
with the Advanced Camera for Surveys (ACS) in the F606W and F814W
bands (V and i), collected on UT 2014 May 15 as part of the Hubble
Frontier Fields (HFF) survey (PI:J.Lotz,
HST-PID:13495).\footnote{\url{http://www.stsci.edu/hst/campaigns/frontier-fields}}
The HFF program is a 3-year director's discretionary initiative that
is collecting 140 orbits of HST imaging (roughly 340 ksec) on six
massive galaxy clusters, plus 6 accompanying parallel fields.  Each
field is observed in 3 optical bands (ACS F435W, F606W and F814W) and
4 infrared (IR) bands (WFC3-IR F105W, F125W, F140W, and F160W),
although the optical and IR imaging campaigns are separated by $\sim$6
months. Abell 2744 was the first cluster observed, with IR imaging
spanning 2013 October-November, and optical imaging from 2014
May-July.  A composite image of the HFF data showing the SN is
presented in Figure~\ref{fig:DiscoveryImage}, \changeR{and the
locations of the cluster center, the SN, and the presumed host galaxy
are given in Table~\ref{tab:Positions}.}  The SN detection was made in
difference images constructed using template imaging of Abell 2744
from HST+ACS observations taken in 2009 (PI:Dupke, HST-PID:11689).

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/host_sed_fit}
\caption{  \label{fig:HostSED}
Best-fit SED template match for the most probable \tomas\ host
galaxy. Using the {\it BPZ} code to match the observed SED, we find
the nearest galaxy to the SN position is a late type galaxy with a
photometric redshift of $z=1.5\pm0.2$.}
\end{center}
\end{figure}

The most probable host galaxy for SN \tomas\ is a faint and diffuse
galaxy immediately to the south-east of the SN location.  With
photometry of the host galaxy collected from the template images, we
fit the spectral energy distribution (SED) using the {\it BPZ} code --
a Bayesian photometric redshift estimator \citep{Benitez:2000}. The
best-fit SED template match is shown in Figure~\ref{fig:HostSED}.
From the {\it BPZ} analysis, we found the host to be most likely an
actively star-forming galaxy at a redshift of $z=1.5\pm0.2$.

Another nearby bright galaxy to the East of the SN has a spectroscopic
redshift of $z=1.742$.  This redshift was determined from a spectrum
taken with the G141 grism of the HST WFC3-IR camera, collected as part
of the Grism Lens Amplified Survey from Space (GLASS, PI:Treu,
PID:13459, \citealt{Treu:2015}). \change{As we will see in Sections~\ref{sec:Spectroscopy}
and \ref{sec:PhotometricClassification}, the SN data are inconsistent
with that redshift, meaning that this is a background
object and therefore has no impact on the SN magnification.}

Upon discovery, HST target-of-opportunity observations were triggered
from the FrontierSN program (PI:Rodney, HST-PID:13386), which aims to
discover and follow transient sources in the HFF cluster and parallel
fields. The FrontierSN observations provided WFC3-IR imaging as well
as spectroscopy using the ACS G800L grism, supplementing the
rapid-cadence optical imaging from HST+ACS already being provided by
the HFF program. \change{The last detections in the IR F105W and F140W
bands came from the direct-imaging component of the GLASS program}.
Difference images for the IR follow-up data were generated using
templates constructed from the HFF WFC3-IR imaging campaign, which
concluded in November, 2013.

@@

Figure~\ref{fig:SpecFit} shows the composite 1-D ACS grism spectrum,
combining all available G800L exposures, overlaid with SN model fits
that will be described below.  The spectrum is largely free of
contamination, because the orientation was chosen to avoid nearby
bright sources and the host galaxy is diffuse and optically faint.
Thus, the SN spectral features can be unambiguously identified, most
notably the red slope of the continuum and a prominent absorption
feature at $\sim$8700\AA.  However, the signal to noise ratio (S/N)
for the host galaxy spectrum was too weak to yield any additional
constraints on the host type or redshift.

\changeG{
Spectra of Type I SNe (including all sub-classes Ia, Ib and Ic) are
dominated by broad absorption features, which cannot in general be
used to directly extract a spectroscopic redshift \citep[see
e.g.][]{Filippenko:1997}.  As described below, we fit template spectra
to the SN \tomas\ data in two steps.}  First we determine a spectral
classification -- and get a preliminary estimate of the redshift and
age -- using the SuperNova IDentification (SNID)
software \citep{Blondin:2007}.  Second, we refine the redshift and age
measurement using a custom Type Ia spectral template matching program.

@@

The SNID program is designed to estimate the type, redshift, and age
of a SN spectrum through cross-correlation matching with a library of
template spectra, \change{using the algorithm of \citet{Tonry:1979}.}
To account for possible distortions in the broad shape of the SN
pseudo-continuum due to dust or instrumental calibration effects, SNID
divides each SED by a smooth cubic spline fit. This effectively
removes the shape of the SN pseudo-continuum to leave behind a flat
SED superimposed with spectral absorption and emission features.  It
is these features which drive the cross-correlation fit, so the SNID
approach is insensitive to the overall color of the SED.  We used v2.0
of the SNID template library, which includes template SEDs covering
all Type Ia and Core Collapse sub-classes, and has recently been
updated with corrections and improvements to the Type Ib/c
templates \citep{Liu:2014}.

@@

To match the SN \tomas\ spectrum we use conservative constraints on
age and redshift: limiting the age to $\pm$5 rest-frame days of peak
brightness and $0.8<z<1.8$, consistent with the SN light curve and
host galaxy photo-$z$.  With these constraints we find that the only
acceptable match is a normal Type Ia SN near $z=1.3$. The best match
has {\it rlap}$=8.7$, using the normal Type Ia SN 2005cf at $z=1.35$
and age=-2.2 rest-frame days before peak.  Within these constraints,
the best non-Ia matches all have {\it rlap}$<2.5$.

@@

\subsection{Spectroscopic Redshift}
\label{sec:SpecRedshift}

To refine the redshift and phase constraints on \tomas, we next fit
the spectrum with a custom spectral matching program that employs a
library of Type Ia SN SEDs.  This library is similar to the Type Ia
spectral set used by SNID, but also includes more recent SNe with
well-observed spectral time series that extend to rest-frame UV
wavelengths (e.g. SN 2011fe and 2014J).  We first use an approach
similar to the SNID algorithm: warping the pseudo-continuum of each
template spectrum by dividing out a 3$^{\rm rd}$-order polynomial to
match the observed SED of SN \tomas.  From this analysis we find
results that are consistent with the SNID fits: a redshift of
$z=1.31\pm0.01$ and a phase of $0\pm3$ rest-frame days. The
best-fitting spectral template is the normal Type Ia SN 2012cg, shown
in the top panel of Figure~\ref{fig:SpecFit}, which has $\chi^2$
per degree of freedom $\nu$ equal to 42.1/87.

Finally, we repeat the fitting, but without any warping of the
templates to account for differences in the continuum shape.  In this
iteration we only allow each template SED to be scaled in flux
coherently at all wavelengths.  As shown in Figure~\ref{fig:SpecFit},
we again find that the \tomas\ SED can be matched by a normal Type Ia
SN (SN 2011fe) at redshift $z=1.31\pm0.01$.  Without the continuum
warping, an alternative fit also arises: the \change{fast-declining
(91bg-like)} Type Ia SN 2011iv
at $z=0.98\pm0.01$. Formally, this match provides a slightly better
fit to the unwarped \tomas\ spectrum, although the fit is notably poor
at $\sim8700$\AA\ where the most significant absorption feature is
found. Furthermore, a redshift $z\sim1$ is at odds with the host
galaxy photo-z ($1.5\pm0.2$), and we will see in the following section
that the photometric data is also incompatible with a Type Ia SN at
$z\sim1.0$.

@@

The final photometric classification probability for SN \tomas\ is
$p({\rm Ia}|{\rm {\bf D}})=1.0$, with the classification probability
from all \CCSN\ sub-classes totaling less than
$10^{-32}$.  \change{Although this Bayesian classification utilizes
the full posterior probability distribution, for illustration we
highlight in Figure~\ref{fig:photoclass} a single best-fit model for
each sub-class.} This demonstrates how the \CCSN\ models fail to
adequately match the observed photometry. In particular, only the Type
Ia model can simultaneously provide an acceptable fit to the
well-sampled rising light curve in F814W and the F814W-F160W color
near peak.

@@

As in \P14, we derive a distance modulus\footnote{We use 'dm' to
indicate the distance modulus to avoid confusion, reserving the symbol
$\mu$ to refer to the lensing magnification. This 'dm' is a standard
distance modulus, defined as ${\rm dm}=5\log_{10}d_L+25$, where $d_L$
is the luminosity distance in Mpc.} \changeG{while ignoring the
effects of gravitational lensing, so that we can later use it to
derive a direct measurment of the magnification.}  From the SALT2 fit
we get

@@

\change{
The total uncertainty in the distance modulus is  
}

@@

\change{
\noindent
The $\sigma_{\rm stat}$ term is the statistical uncertainty, which
encapsulates uncertainties from the data and the model, and
$\sigma_{\rm int}$ accounts for the remaining {\it unmodeled} scatter.
This latter term is derived by finding the amount of additional
distance modulus scatter that needs to be added to a Type Ia SN
population to get $\chi^2$ per degree of freedom equal to 1 for a
fiducial cosmological model fit to the SN Ia Hubble diagram (distance
modulus vs redshift).  Thus, $\sigma_{\rm int}$ is designed to
account for any unknown sources of scatter in the Type Ia SN
population, including unidentified errors in the data analysis as well
as the natural scatter in intrinsic Type Ia SN luminosities.  The value of 
$\sigma_{\rm int}$ may be expected to vary as a function of
redshift and also from survey to survey.  
}

>>>>

\begin{abstract}
SN \tomas\ is a Type Ia Supernova (\SNIa) discovered at
\change{$z=1.3457\pm0.0001$} behind the galaxy cluster Abell 2744 ($z=0.308$). This
SN has a projected separation from the cluster core of
$\sim$40\arcsec, closer to the critical lensing line than any previous
cluster-lensed Type Ia SN.  As such, it provides a valuable
opportunity to confront gravitational lens models of galaxy clusters
with the direct measurement of an absolute magnification at the edge
of the strong-lensing region.  We derive a tightly constrained measure
of the peak apparent magnitude, corrected for light curve shape and
exinction.  In a cosmology-independent analysis, we find that \tomas\
is \change{$0.77\pm0.15$} magnitudes brighter than unlensed Type Ia
SNe at similar redshift, implying a lensing magnification
of \change{$\mu_{\rm obs}=2.03\pm0.29$}.  We compare this direct
measurement of the magnification against predicted magnifications
from 14 lens models, representing a broad range of
methodologies and incorporating a variety of strong- and weak-lensing
constraints.  The models are collectively fairly accurate: 6 of the
models deliver median magnifications that are consistent with the
measured $\mu$ to within the 1$\sigma$ uncertainties.  However, the
models are systematically biased to higher values: \change{all but
one} of the tested models yield a median magnification higher than the
observed $\mu$, and \change{6 of the models are $\geq1.5\sigma$}
above the measured magnification.  We evaluate possible causes for
this mild bias, considering both modeling errors and
misinterpretation of the SN data. Although we find no single
explanation that can resolve this, we \change{ highlight several ways
that lens models can deliver more realistic uncertainties, such as
incorporating errors from assumed cosmological parameters and allowing
greater flexibility in the form of the cluster mass distribution.  A
more robust evaluation of systematic biases could be achieved with
further cooperative lens model comparisons using a larger sample of
lensed Type Ia SNe and cluster lensing simulations.}
\end{abstract}

\keywords{ supernovae: general, supernovae: individual: HFF14Tom, galaxies: clusters: general, galaxies: clusters: individual: Abell 2744, gravitational lensing: strong, gravitational lensing: weak  }

@@

Galaxy clusters can be used as cosmic telescopes to magnify distant
background objects through gravitational lensing, which can
substantially increase the reach of deep imaging surveys.  The lensing
magnification enables the study of objects that would otherwise be
unobservable because they are either intrinsically
faint \citep[e.g.][]{Schenker:2012,Alavi:2014} or extremely
distant \citep[e.g.][]{Franx:1997,Ellis:2001,Hu:2002,Kneib:2004,Richard:2006,Richard:2008,Bouwens:2009a,Maizy:2010,Zheng:2012,Coe:2013,Bouwens:2014,Zitrin:2014b}.
Background galaxies are also {\it spatially} magnified, allowing for
studies of the internal structure of galaxies in the early universe
with resolutions of $\sim$100
pc \citep[e.g.]{Stark:2008,Jones:2010,Yuan:2011,Livermore:2015}.

Gravitational lensing can also provide a powerful window onto the {\it
transient} sky through an appropriately cadenced imaging survey.  The
flux magnification from strong-lensing clusters is especially valuable
for the study of $z>1.5$ supernovae
(SNe) \citep[e.g.]{Kovner:1988,Kolatt:1998,Sullivan:2000,Saini:2000,Gunnarsson:2003,Goobar:2009,Postman:2012},
which are still extremely difficult to characterize in unlensed
fields \citep[e.g.]{Riess:2001,Riess:2007,Suzuki:2012,Rodney:2012,Rubin:2013,Jones:2013}.

In the case of lensed SNe, we can also reverse the experimental setup:
instead of using strong-lensing clusters to study distant SNe, we can
use the SNe as tools for examining the lenses \citep{Riehm:2011}.  The
most valuable transients for testing and improving cluster lens models
would be {\it strongly lensed} SNe that are resolved into multiple
images \citep{Holz:2001,Oguri:2003}.  Transients that are lensed into
multiple images can also become cosmological tools, as the measurement
of time delays between the images can provide cosmographic information
to constrain the Hubble parameter \citep{Refsdal:1964} and other
cosmological parameters \citep{Linder:2011}.  We have recently
observed the first example of a multiply-imaged
SN \citep{Kelly:2015}, and expect to detect the reappearance of this
object (called ``SN Refsdal'') within 5 years, delivering a precise
test of lens model predictions \citep{Oguri:2015,Sharon:2015}.
Although detections of such objects are currently very
unlikely \citep{Li:2012}, they will become much more common in the
next decade \citep{Coe:2009,Dobke:2009}, and may be developed into an
important new cosmological tool.

In addition to time delays from multiply-imaged SNe, we can also
put cluster mass models to the test with the much more common category
of lensed Type Ia SNe.  
\citet[][hereafter P14]{Patel:2014} and \citet{Nordin:2014} presented
independent analyses of three lensed SNe, of which at least 2 are
securely classified as Type Ia SNe -- all found in the Cluster Lensing
and Supernova survey with Hubble (CLASH, PI:Postman, HST Program ID
12068, \citealt{Postman:2012}).  Both groups demonstrated that these
standard candles can be used to provide accurate and precise
measurements of the true absolute magnification along a random sight
line through the cluster.  Although in these cases the SNe were used
to {\it test} the cluster mass models, one could in principle
incorporate the measured magnifications of Type Ia SNe into the
cluster as additional model constraints.  In that role, Type Ia SNe
have the particular value that they can be found anywhere in the
cluster field.  Thus, they can deliver model constraints in regions of
``middle distance'' from the cluster core, where both strong- and
weak-lensing constraints are unavailable.

One of the key values in observing standard candles behind
gravitational lenses is in addressing the problem of the
mass-sheet degeneracy \citep{Falco:1985,Bradac:2004}.  This
degeneracy arises because one can introduce into a lens model an
unassociated sheet of uniform mass in front of or behind the
lens, without disturbing the primary observable quantities.  For
example, take a lens model with a given surface mass density $\kappa$,
and then transform the surface mass density to
$\kappa^{\prime}=(1-\lambda)\kappa+\lambda$ for any arbitrary value
$\lambda$.  Both the $\kappa$ and $\kappa^{\prime}$ models will
produce exactly the same values for all positional and shear
constraints from strong and weak lensing.  When lensed background
sources are available across a wide range of redshifts (as is the case
for the Abell 2744 cluster discussed here), it should in principle be
possible to account for the simplest form of a single mass sheet
within that redshift range.  However, there are more complex versions
of positional constraint
degeneracies \citep{Schneider:2014,Liesenborgs:2012}.  Such
degeneracies do not extend to the {\it absolute} magnification of a
background source's flux and size.  Therefore, an absolute measurement
of magnification from a standard candle or a standard ruler
\citep{Sonnenfeld:2011}  can break these fundamental degeneracies
\citep{Holz:2001}.

In Section~\ref{sec:DiscoveryAndFollowup} we present the discovery and
follow-up observations of SN \tomas\ at $z=1.3457$, discovered behind the
galaxy cluster Abell 2744.  Sections~\ref{sec:Spectroscopy}
and \ref{sec:PhotometricClassification} describe the spectroscopy and
photometry of this SN, leading to a classification of the object as a
normal Type Ia SN.  In Section~\ref{sec:LightCurveFitting} we make a
direct measurement of the magnification of this source due to
gravitational lensing, and compare to predictions from lens models.
Finally, Section~\ref{sec:Discussion} discusses the tension between
our magnification measurement and the lens models, with implications
for the systematic error budget in magnification estimates for other
high-$z$ lensed sources.

@@

SN \tomas\ was discovered in Hubble Space Telescop (\HST) observations
with the Advanced Camera for Surveys (ACS) in the F606W and F814W
bands (V and i), collected on UT 2014 May 15 as part of the Hubble
Frontier Fields (HFF) survey (PI:J.Lotz,
HST-PID:13495).\footnote{\url{http://www.stsci.edu/hst/campaigns/frontier-fields}}
The HFF program is a 3-year director's discretionary initiative that
is collecting 140 orbits of HST imaging (roughly 340 ksec) on six
massive galaxy clusters, plus 6 accompanying parallel fields.  Each
field is observed in 3 optical bands (ACS F435W, F606W and F814W) and
4 infrared (IR) bands (WFC3-IR F105W, F125W, F140W, and F160W),
although the optical and IR imaging campaigns are separated by $\sim$6
months. Abell 2744 was the first cluster observed, with IR imaging
spanning 2013 October-November, and optical imaging from 2014
May-July.  A composite image of the HFF data showing the SN is
presented in Figure~\ref{fig:DiscoveryImage}, and the
locations of the cluster center, the SN, and the presumed host galaxy
are given in Table~\ref{tab:Positions}.  The SN detection was made in
difference images constructed using template imaging of Abell 2744
from HST+ACS observations taken in 2009 (PI:Dupke, HST-PID:11689).

@@

%\begin{figure}
%\begin{center}
%\includegraphics[width=\columnwidth]{FIG/host_sed_fit}
%\caption{  \label{fig:HostSED}
%Best-fit SED template match for the most probable \tomas\ host
%galaxy. Using the {\it BPZ} code to match the observed SED, we find
%the nearest galaxy to the SN position is a late type galaxy with a
%photometric redshift of $z=1.5\pm0.2$.}
%\end{center}
%\end{figure}

The most probable host galaxy for SN \tomas\ is a faint and diffuse
galaxy immediately to the south-east of the SN location.  With
photometry of the host galaxy collected from the template images, we
fit the spectral energy distribution (SED) using the {\it BPZ} code --
a Bayesian photometric redshift estimator \citep{Benitez:2000}. 
%The best-fit SED template match is shown in Figure~\ref{fig:HostSED}.
From the {\it BPZ} analysis, we found the host to be most likely an
actively star-forming galaxy at a redshift of
$z=1.5\pm0.2$.  \change{This photo-z was subsequently supported by a
spectroscopic redshift of $z=1.3457\pm0.0001$, based on emission lines
consistent with \textcolor{red}{??? the [OII] $\lambda\lambda$ 3726-3729 \AA\ doublet.
???}}  Another nearby bright galaxy to the East of the SN has a
spectroscopic redshift of $z=1.742$.  This redshift was determined
from a spectrum taken with the G141 grism of the HST WFC3-IR camera,
collected as part of the Grism Lens Amplified Survey from Space
(GLASS, PI:Treu, PID:13459, \citealt{Treu:2015}). As we will see in
Sections~\ref{sec:Spectroscopy}
and \ref{sec:PhotometricClassification}, the SN data are \change{consistent
with $z=1.3457$,} meaning that the latter galaxy is a background object and
therefore has no impact on the SN magnification.

Upon discovery, HST target-of-opportunity observations were triggered
from the FrontierSN program (PI:Rodney, HST-PID:13386), which aims to
discover and follow transient sources in the HFF cluster and parallel
fields. The FrontierSN observations provided WFC3-IR imaging as well
as spectroscopy of the SN itself using the ACS G800L grism,
supplementing the rapid-cadence optical imaging from HST+ACS already
being provided by the HFF program. The last detections in the IR F105W
and F140W bands came from the direct-imaging component of the GLASS
program.  Difference images for the IR follow-up data were generated
using templates constructed from the HFF WFC3-IR imaging campaign,
which concluded in November, 2013.

@@

Figure~\ref{fig:SpecFit} shows the composite 1-D ACS grism spectrum,
combining all available G800L exposures, overlaid with SN model fits
that will be described below.  The spectrum is largely free of
contamination, because the orientation was chosen to avoid nearby
bright sources and the host galaxy is diffuse and optically faint.
Thus, the SN spectral features can be unambiguously identified, most
notably the red slope of the continuum and a prominent absorption
feature at $\sim$8700\AA.  
%However, the signal to noise ratio (S/N)
%for the host galaxy spectrum was too weak to yield any additional
%constraints on the host type or redshift.

Spectra of Type I SNe (including all sub-classes Ia, Ib and Ic) are
dominated by broad absorption features, which cannot in general be
used to directly extract a spectroscopic redshift \citep[see
e.g.][]{Filippenko:1997}.  As described below, we fit template spectra
to the SN \tomas\ data in two steps.  First we determine a spectral
classification -- and get a preliminary estimate of the redshift and
age -- using the SuperNova IDentification (SNID)
software \citep{Blondin:2007}.  Second, we refine the redshift and age
measurement using a custom Type Ia spectral template matching program.

@@

The SNID program is designed to estimate the type, redshift, and age
of a SN spectrum through cross-correlation matching with a library of
template spectra, using the algorithm of \citet{Tonry:1979}.
To account for possible distortions in the broad shape of the SN
pseudo-continuum due to dust or instrumental calibration effects, SNID
divides each SED by a smooth cubic spline fit. This effectively
removes the shape of the SN pseudo-continuum to leave behind a flat
SED superimposed with spectral absorption and emission features.  It
is these features which drive the cross-correlation fit, so the SNID
approach is insensitive to the overall color of the SED.  We used v2.0
of the SNID template library, which includes template SEDs covering
all Type Ia and Core Collapse sub-classes, and has recently been
updated with corrections and improvements to the Type Ib/c
templates \citep{Liu:2014}.

@@

To match the SN \tomas\ spectrum we use conservative constraints on
age and redshift: limiting the age to $\pm$5 rest-frame days from peak
brightness and $0.8<z<1.8$, consistent with the SN light curve and
the two plausible host galaxies.  With these constraints we find that the only
acceptable match is a normal Type Ia SN near $z=1.3$. The best match
has {\it rlap}$=8.7$, using the normal Type Ia SN 2005cf at $z=1.35$
and age=-2.2 rest-frame days before peak.  In contrast, 
the best non-Ia matches all have {\it rlap}$<2.5$.

@@

\subsection{Spectral Fitting with UV Type Ia Templates}
\label{sec:SpecRedshift}

To refine the redshift and phase constraints on \tomas, we next fit
the spectrum with a custom spectral matching program that employs a
library of Type Ia SN SEDs.  This library is similar to the Type Ia
spectral set used by SNID, but also includes more recent SNe with
well-observed spectral time series that extend to rest-frame UV
wavelengths (e.g. SN 2011fe and 2014J).  We first use an approach
similar to the SNID algorithm: warping the pseudo-continuum of each
template spectrum by dividing out a 3$^{\rm rd}$-order polynomial to
match the observed SED of SN \tomas.  From this analysis we find
results that are consistent with the SNID fits \change{and the host
galaxy redshift}: $z=1.31\pm0.03$ and a phase of $0\pm3$ rest-frame
days. The best-fitting spectral template is the normal Type Ia SN
2012cg, shown in the top panel of Figure~\ref{fig:SpecFit}, which has
$\chi^2$ per degree of freedom $\nu$ equal to 42.1/87.  \change{If we
fix the redshift at $z=1.3457$ then we find the best fit is from the
normal Type Ia SN 2014J, with $\chi^2/\nu=51.8/87$.}

Next, we repeat the fitting, but without any warping of the templates
to account for differences in the continuum shape.  In this iteration
we only allow each template SED to be scaled in flux coherently at all
wavelengths.  As shown in Figure~\ref{fig:SpecFit}, we again find that
the \tomas\ SED can be matched by a normal Type Ia SN (SN 2011fe) at
redshift $z=1.31\pm0.02$.  \change{Fixing the redshift at the host redshift of
$z=1.3457$, we also find the best match is from SN 2011fe, with
$\chi^2/\nu=133.7/91$. }

Without the continuum warping, an alternative
fit also arises: the fast-declining (91bg-like) Type Ia SN 2011iv at
$z=0.98\pm0.01$. Formally, this match provides a slightly better fit
to the unwarped \tomas\ spectrum, although the fit is notably poor at
$\sim8700$\AA\ where the most significant absorption feature is
found. Furthermore, a redshift $z\sim1$ is at odds with the host
galaxy photo-z ($1.5\pm0.2$), and we will see in the following section
that the photometric data is also incompatible with a Type Ia SN at
$z\sim1.0$.

@@

The final photometric classification probability for SN \tomas\ is
$p({\rm Ia}|{\rm {\bf D}})=1.0$, with the classification probability
from all \CCSN\ sub-classes totaling less than
$10^{-32}$.  Although this Bayesian classification utilizes
the full posterior probability distribution, for illustration we
highlight in Figure~\ref{fig:photoclass} a single best-fit model for
each sub-class. This demonstrates how the \CCSN\ models fail to
adequately match the observed photometry. In particular, only the Type
Ia model can simultaneously provide an acceptable fit to the
well-sampled rising light curve in F814W and the F814W-F160W color
near peak.

@@

As in \P14, we derive a distance modulus\footnote{We use 'dm' to
indicate the distance modulus to avoid confusion, reserving the symbol
$\mu$ to refer to the lensing magnification. This 'dm' is a standard
distance modulus, defined as ${\rm dm}=5\log_{10}d_L+25$, where $d_L$
is the luminosity distance in Mpc.} while ignoring the
effects of gravitational lensing, so that we can later use it to
derive a direct measurment of the magnification.  From the SALT2 fit
we get

@@

The total uncertainty in the distance modulus is

@@

\noindent
The $\sigma_{\rm stat}$ term is the statistical uncertainty, which
encapsulates uncertainties from the data and the model, and
$\sigma_{\rm int}$ accounts for the remaining {\it unmodeled} scatter.
This latter term is derived by finding the amount of additional
distance modulus scatter that needs to be added to a Type Ia SN
population to get $\chi^2$ per degree of freedom equal to 1 for a
fiducial cosmological model fit to the SN Ia Hubble diagram (distance
modulus vs redshift).  Thus, $\sigma_{\rm int}$ is designed to
account for any unknown sources of scatter in the Type Ia SN
population, including unidentified errors in the data analysis as well
as the natural scatter in intrinsic Type Ia SN luminosities.  The value of 
$\sigma_{\rm int}$ may be expected to vary as a function of
redshift and also from survey to survey.
